---
title: "Final Model Project"
author: "Juliette Verstaen and Seleni Cruz"
date: "5/12/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(magrittr)
library(tidyverse)
library(deSolve)
library(dplyr)
library(zeallot)

```

###Going through JC stuff 

Run Model to Equilibrium
```{r}

source("R/wrapper.R")
source("R/predprey.R")
source("R/extract.R")

# First call of the model to reach equilibrium
start <- wrapper_model(hx = 0, hp = 0, t0=0, t1 = 30, X0 = 25, P0 = 25)

# Extract final conditions and start trophic downgrading
c(X1, P1) %<-%  extract_state_vars(start)
harvest_P <- wrapper_model(hx = 0, t0 = 30, t1 = 70, X0 = X1, P0 = P1)

# Extract final conditions and start fishing for prey too
c(X2, P2) %<-% extract_state_vars(harvest_P)
harvest_X <- wrapper_model(t0 = 70, t1 = 120, X0 = X2, P0 = P2)

```

Syncronous Recovery
What about happen if fishing was completly halted
```{r}

# Stop fishing and let both species recover
c(X3, P3) %<-% extract_state_vars(harvest_X)

stop_harvest <- wrapper_model(hx = 0, hp = 0, t0 = 120, t1 = 200, X0 = X3, P0 = P3)
# create ggplot object
both <- rbind(start, harvest_P, harvest_X, stop_harvest) %>% ggplot(aes(x = Time, y = Abundance)) +
geom_line(aes(color = Organism), size = 1) + scale_color_brewer(palette = "Set1") + theme(legend.position = "none")

both

```

### Adding on MPA management


```{r}

library(magrittr)
library(tidyverse)
library(deSolve)
library(dplyr)
library(zeallot)

source("R/wrapperTEST.R")
source("R/mpa_var_setup.R")
source("R/run_mpa.R")
source("R/predprey.R")
source("R/extract.R")


# these parameters are on the outside for testing:
mrate_X = 0.5
mrate_P = 0.5
ncells=10
MPA_width=0
hx = 0.65
hp = 0.32
rx = 1.2
Kx = 100
ax = 0.03
c = 0.05
ay = 0.03
dp = 0.25
Kp = 25
t0 = 0
t1 = 40
t0=0
t1 = 30
X0 = 25
P0 = 25

# First call of the model to reach equilibrium
# this should be the same results as the above part with JC becuase there is no MPA
# this is also where there is an issue I can't figure out
start_test <- wrapper_modelTEST(hx = 0, hp = 0, t0=0, t1 = 30, X0 = 25, P0 = 25, MPA_width=0)

# Extract final conditions 
c(X1, P1) %<-%  extract_state_vars(start)

# try it with an MPA width
testmpa <- wrapper_modelTEST(t0=0, t1 = 30, X0=X1, P0=P1, MPA_width=4)

```

**Species we will use:**
Predator (P) -> goliath grouper
Prey (X) -> yellow snapper 
Unessential Prey (Y) -> parrotfish 

**Parameters:**
goliah grouper growth rate (c) = 0.2 +- 0.03
movement = 0.5 for both

FINISHING FILLING THESE OUT. we looked some up already but we should probs agree on what we want them to be

**Issues:**
Error in func(time, state, parms, ...) : unused arguments (state, parms)

**Things to rememeber/clean up:**
1. go back through and make sure the naming convention of things make sense. if it's saved as vec, is it really a vector
2. meta data

**Sensitivity Analysis:**
We want to do it for c, r, and hx, hp



