---
title: "Final Report: Recovery and synchronized management of fisheries"
author: "Seleni Cruz and Juliette Verstean"
date: "June 3, 2019"
output: html_document
bibliography: ["references.bib"]
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(deSolve)
library(ggplot2)
library(sensitivity)
library(foreach)
source("R/predprey.R")
source("R/wrapper.R")
source("R/sensitivity.R")
source("R/mpa.R")
```

Selecting generalist predator and prey
For the simulations we used general predator and prey charactieristics, as used in @samhouriRapidDirectRecoveries2017 (Table 1). 

```{r, echo= FALSE}

Varible <- list("P", "X", "Y", "rx", "dp", "kp", "kx", "ax", "ay", "hp", "hx", "c")

Description <- list("Predator", "Focal prey", "Non-dynamic prey", " Growth rate prey", "Predator mortality rate", "Predator carrying capacity", "Prey carrying capacity", "Predation rate on focal prey", "Predation rate on non-dynamic prey", "Predator harvest rate", "Focal prey harvest rate", "Predator conversion rate of prey") 
Values <- list(50, 30, 500, 1, 0.25, 25, 100, 0.03, 0.03, 0.325, 0.65, 0.05)

Units <- list("# individuals", "# individuals", "# individuals", "y^-1", "y^-1", "# individuals",  "# individuals", "# individuals^-1*y^-1", "# individuals^-1*y^-1","y^-1", "y^-1", "prey/predator"  )

parameters <- cbind(Varible, Description, Values, Units) %>%data.frame()

knitr::kable(parameters, align= 'l', caption= "List of parameters for predator-prey model")
```


Simulating trophic downgrading

We used the predator prey and the wrapper function to first simulate trophic downgrading (Figure 1). Three steps included:
1. Using initial predator and prey populations (X0 and P0) we ran the model to equilibrium for the first 30 timesteps
2. After equilibrium, we began harvesting the predator first. Simulating a scenario where fishers would first target these fisheries. 
3. After the population has been depleted, we simulated harvesting the prey. 



```{r echo=FALSE}
start <- run_model(hx = 0, hp = 0, t1 = 30, X0 = 30, P0 = 50)

X1 <- tail(start$X, 1)
P1 <- tail(start$P, 1)
harvest_P <- run_model(hx = 0, t0 = 30, t1 = 60, X0 = X1, P0 = P1)

X2 <- tail(harvest_P$X, 1)
P2 <- tail(harvest_P$P, 1)
harvest_X <- run_model(t0 = 60, t1 = 90, X0 = X2, P0 = P2)

X3 <- tail(harvest_X$X, 1)
P3 <- tail(harvest_X$P, 1)

down <- rbind(start, harvest_P, harvest_X) %>% select(Time, X, P)

scen1 <- rbind(start, harvest_P, harvest_X)%>%
   select(Time, X, P)%>%
  tidyr::gather(key="Organism", value="Abundance", -Time)

ggplot(scen1, aes(x=Time, y=Abundance, color=Organism))+geom_line(size=3)+
  scale_color_discrete(" ", labels=c("Predator", "Prey"))+
  geom_vline(xintercept=30, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=60, linetype="dashed", color="black", size=1)+
  labs(subtitle = "Figure 1: Simulating trophic downgrading")+
  annotate("text", x = 0, y = 65, label = "(1)")+
  annotate("text", x = 35, y = 65, label = "(2)")+
  annotate("text", x = 65, y = 65, label = "(3)")+
  theme_classic()

```

First management interventon: first removing harvesting from predator and then prey 

After simulating trophic downgrading we simulated single species management interventions, protection of predator first and then prey (Figure 2). 
```{r echo=FALSE}
stop_harvest_P <- run_model(hp = 0, t0 = 90, t1 = 120, X0 = X3, P0 = P3)

X4 <- tail(stop_harvest_P$X, 1)
P4 <- tail(stop_harvest_P$P, 1)

stop_harvest_X <- run_model(hx = 0, hp = 0, t0 = 120, t1 = 200, X0 = X4, P0 = P4)

scen1 <- rbind(start, harvest_P, harvest_X, stop_harvest_P, stop_harvest_X)%>%
   select(Time, X, P)%>%
  tidyr::gather(key="Organism", value="Population", -Time)

sum1 <- rbind(stop_harvest_P, stop_harvest_X)%>% 
  summarise_at(vars(X, P, H_X, H_P), sum) %>% mutate(Name= "Pred")
  

ggplot(scen1, aes(x=Time, y=Population, color=Organism))+geom_line(size=3)+
  scale_color_discrete(" ", labels=c("Predator", "Prey"))+
   labs(subtitle = "Figure 2: Single species management, predator then prey")+
  geom_vline(xintercept=30, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=60, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=90, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=120, linetype="dashed", color="black", size=1)+
  annotate("text", x = 0, y = 65, label = "(1)")+
  annotate("text", x = 35, y = 65, label = "(2)")+
  annotate("text", x = 65, y = 65, label = "(3)")+
  annotate("text", x = 95, y = 65, label = "(4)")+
  annotate("text", x = 125, y = 65, label = "(5)")+
  theme_classic()

#ggsave("Pred_Prey.png", width = 8, height = 3, dpi=400)

```

Second management interventon: first removing harvesting from prey and then predator 

The second management scenrio simulated the protection of prey first and then predator (Figure 3). 
```{r echo=FALSE}
stop_harvest_X <- run_model(hx = 0, t0 = 90, t1 = 120, X0 = X3, P0 = P3)

X4 <- tail(stop_harvest_X$X, 1)
P4 <- tail(stop_harvest_X$P, 1)

stop_harvest_P <- run_model(hp = 0, hx = 0, t0 = 120, t1 = 200, X0 = X4, P0 = P4)

scen1 <- rbind(start, harvest_P, harvest_X, stop_harvest_X, stop_harvest_P)%>%
   select(Time, X, P)%>%
  tidyr::gather(key="Organism", value="Population", -Time)

sum2 <- rbind(stop_harvest_X, stop_harvest_P)%>% 
  summarise_at(vars(X, P, H_X, H_P), sum) %>% mutate(Name= "Prey")

ggplot(scen1, aes(x=Time, y=Population, color=Organism))+geom_line(size=3)+
  scale_color_discrete(" ", labels=c("Predator", "Prey"))+
    labs(subtitle = "Figure 3: Single species management, prey then predator")+
 geom_vline(xintercept=30, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=60, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=90, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=120, linetype="dashed", color="black", size=1)+
  annotate("text", x = 0, y = 70, label = "(1)")+
  annotate("text", x = 35, y = 70, label = "(2)")+
  annotate("text", x = 65, y = 70, label = "(3)")+
  annotate("text", x = 95, y = 70, label = "(4)")+
  annotate("text", x = 125, y = 70, label = "(5)")+
  theme_classic()

#ggsave("Prey_pred.png", width = 8, height = 3, dpi=400)

```

Third management intervention: simulating syncronized closure to both predator and prey

We simulated combined closures in the form of marine proteted areas (MPAs) by removing harvesting. For example a marine protected area that covers 50% of the area is simulated by removing 50% of the haverst rate for both predator and prey. While we initally wanted to make this part of the model spatial, area represented by a series of vectors, we were not able to do so using ODE since inital values of ODE cannot be vectors but rather single values. 

MPAs were simulated from 0 to 100% in 10% increments. Figure 4 and 5 present 50% and 100% closures, respectively. 
```{r echo=FALSE}

MPA <- MPA_sim(hp=0.325 , hx=0.65, X3=X3, P3=P3)

fifty <- MPA%>% filter(MPA==0.5)%>%
  select(Time, X, P)%>%rbind(down)%>%
  tidyr::gather(key="Organism", value="Population", -Time)

ggplot(fifty, aes(x=Time, y=Population, color=Organism))+geom_line(size=3)+
  scale_color_discrete(" ", labels=c("Predator", "Prey"))+
  labs(subtitle = "Figure 4: Synchronized management, 50% MPA closure")+
 geom_vline(xintercept=30, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=60, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=90, linetype="dashed", color="black", size=1)+
  annotate("text", x = 0, y = 70, label = "(1)")+
  annotate("text", x = 35, y = 70, label = "(2)")+
  annotate("text", x = 65, y = 70, label = "(3)")+
  annotate("text", x = 95, y = 70, label = "(4)")+
  theme_classic()

#ggsave("MPA_50.png", width = 8, height = 3, dpi=900)  


all <- MPA%>% filter(MPA==1)%>%
  select(Time, X, P)%>%rbind(down)%>%
  tidyr::gather(key="Organism", value="Population", -Time)

ggplot(all, aes(x=Time, y=Population, color=Organism))+geom_line(size=3)+
 scale_color_discrete(" ", labels=c("Predator", "Prey"))+
    labs(subtitle = "Figure 5: Synchronized management, 100% MPA closure")+
 geom_vline(xintercept=30, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=60, linetype="dashed", color="black", size=1)+
  geom_vline(xintercept=90, linetype="dashed", color="black", size=1)+
  annotate("text", x = 0, y = 70, label = "(1)")+
  annotate("text", x = 35, y = 70, label = "(2)")+
  annotate("text", x = 65, y = 70, label = "(3)")+
  annotate("text", x = 95, y = 70, label = "(4)")+
  theme_classic()

#ggsave("MPA_100.png", width = 8, height = 3, dpi=400)  


```

Comparison of scenarios 

We evaluated the differeces in harvest and population of predator and prey in all scenarios. Management scenario 1 (protection of predator first then prey) had the lowest populations abundance and scenario 2 (protection of prey first and predators after) had the highest population (Figure 8). 

We also compared their productivity (harvest), management scenario 1 and 2 performed most poorly when compared to scenarios with syncronized closures. All MPA scenarios resulted in more productivity over the projected time. 

```{r echo=FALSE}
sum3 <- MPA%>% group_by(MPA)%>% 
  summarise_at(vars(X, P, H_X, H_P), sum) %>% rename(Name= MPA)%>%
  rbind(sum1, sum2)%>%  
  tidyr::gather(key = "Organism", value = "Population", X, P, -Name) %>%
  tidyr::gather(key = "Org", value = "Harvest", H_X, H_P, -Name) 

ggplot(sum3, aes(x = Name, y = Population, fill = Organism)) +
  geom_bar(stat = "identity", position = position_stack())+
  labs(subtitle = "Figure 6: Adundance of predator and prey")+
  labs(x="Scenarios")+
  scale_fill_discrete(" ", labels=c("Predator", "Prey"))+
  theme_classic()
   
#ggsave("population.png", width = 8, height = 3, dpi=400)  

ggplot(sum3, aes(x = Name, y = Harvest, fill = Organism)) +
  geom_bar(stat = "identity", position = position_stack())+
   labs(subtitle = "Figure 7: Harvest of predator and prey")+
  labs(x="Scenarios")+
  scale_fill_discrete(" ", labels=c("Predator", "Prey"))+
  theme_classic()
   
#ggsave("harvest.png", width = 8, height = 3, dpi=400)  

```

Sensitivity 

We also conduted sensitivity analysis of the model parameters using Sobol for management scenario 1. When we ran the analyis with all variables, the model was most sensitive to initial population of both predator and prey (X0 and P0). When we removed these variables to test for sesnitivity for other parameters, Sobol returned NAN values.
```{r echo=FALSE}
#Sensitivity of ll variables 
nsample <- 200
hx <- 0.65
hp <- 0.325
c <- 0.05
rx <- 1 
dp <- 0.25
ax <- 0.03
ay <- 0.03
ps1 <- cbind.data.frame(hx=rnorm(mean=hx, sd=0.2, n=nsample), hp=rnorm(mean=hp, sd=0.2, n=nsample), 
                       c=rnorm(mean = c, sd= 0.1, n=nsample), rx=rnorm(mean=rx, sd=0.2, n=nsample),
                       dp= rnorm(mean=dp, sd=0.1, n=nsample), X0=runif(min=15, max=100, n=nsample),
                       P0=runif(min=15, max=100, n=nsample), ax= rnorm(mean=ax, sd=0.05, n=nsample),
                       ay= rnorm(mean=ay, sd=0.05, n=nsample), Kx=runif(min=75, max=200, n=nsample),
                       Kp=runif(min=20, max=80, n=nsample), Y=runif(min=300, max=600, n=nsample)  )
ps2 <-  cbind.data.frame(hx=rnorm(mean=hx, sd=0.2, n=nsample), hp=rnorm(mean=hp, sd=0.2, n=nsample), 
                       c=rnorm(mean = c, sd= 0.1, n=nsample), rx=rnorm(mean=rx, sd=0.2, n=nsample),
                       dp= rnorm(mean=dp, sd=0.1, n=nsample), X0=runif(min=15, max=100, n=nsample),
                       P0=runif(min=15, max=100, n=nsample), ax= rnorm(mean=ax, sd=0.05, n=nsample),
                       ay= rnorm(mean=ay, sd=0.05, n=nsample), Kx=runif(min=75, max=200, n=nsample),
                       Kp=runif(min=20, max=80, n=nsample), Y=runif(min=300, max=600, n=nsample)  )
sens_PX <- sobol2007(model = NULL, ps1, ps2)
nsim <- nrow(sens_PX$X)

results <- foreach(i = 1:nsim, .combine="rbind") %do% {
  scen1_sensitivity(hx=as.numeric(sens_PX$X[i,"hx"]), 
        hp=as.numeric(sens_PX$X[i,"hp"]), 
        c=as.numeric(sens_PX$X[i,"c"]), 
        rx=as.numeric(sens_PX$X[i,"rx"]), 
        dp=as.numeric(sens_PX$X[i, "dp"]), 
        X0=as.numeric(sens_PX$X[i,"X0"]), 
        P0=as.numeric(sens_PX$X[i,"P0"]), 
        ax=as.numeric(sens_PX$X[i,"ax"]), 
        ay=as.numeric(sens_PX$X[i,"ay"]),
        Kx=as.numeric(sens_PX$X[i,"Kx"]),
        Kp=as.numeric(sens_PX$X[i,"Kp"]),
        Y=as.numeric(sens_PX$X[i,"Y"]))
  }
sens_PX <- tell(sens_PX, results$X)
# loot at results
sens_PX$S
sens_PX$T
sens_PX <- tell(sens_PX, results$P)
# loot at results
sens_PX$S
sens_PX$T
```


```{r echo=FALSE}
nsample <- 200


ps1 <- cbind.data.frame(hx=rnorm(mean=0.65, sd=0.2, n=nsample), hp=rnorm(mean=0.325, sd=0.2, n=nsample), 
                       c=rnorm(mean = 0.05, sd= 0.1, n=nsample), rx=rnorm(mean=1, sd=0.2, n=nsample),
                       dp= rnorm(mean=0.25, sd=0.1, n=nsample), ax= rnorm(mean=0.03, sd=0.05, n=nsample),
                       ay= rnorm(mean=0.03, sd=0.05, n=nsample))

ps2 <-  cbind.data.frame(hx=rnorm(mean=0.65, sd=0.2, n=nsample), hp=rnorm(mean=0.325, sd=0.2, n=nsample), 
                       c=rnorm(mean = 0.05, sd= 0.1, n=nsample), rx=rnorm(mean=1, sd=0.2, n=nsample),
                       dp= rnorm(mean=0.25, sd=0.1, n=nsample), ax= rnorm(mean=0.03, sd=0.05, n=nsample),
                       ay= rnorm(mean=0.03, sd=0.05, n=nsample))

sens_PX <- sobol2007(model = NULL, ps1, ps2)

nsim <- nrow(sens_PX$X)

results <- foreach(i = 1:nsim, .combine="rbind") %do% {
  scen1_sensitivity(hx=as.numeric(sens_PX$X[i,"hx"]), 
                    hp=as.numeric(sens_PX$X[i,"hp"]), 
                    c=as.numeric(sens_PX$X[i,"c"]), 
                    rx=as.numeric(sens_PX$X[i,"rx"]), 
                    dp=as.numeric(sens_PX$X[i, "dp"]),
                    ax=as.numeric(sens_PX$X[i,"ax"]), 
                    ay=as.numeric(sens_PX$X[i,"ay"]))
  }


sens_PX <- sensitivity::tell(sens_PX, results$X)
# look at results
sens_PX$S
sens_PX$T

sens_PX <- tell(sens_PX, results$P)
# look at results
sens_PX$S
sens_PX$T
```

Key take aways   
1. Predator first then prey recovery is still widely used in fisheries management.    
2. Our results show that in some cases this is the least optimal management intervention to achieve fisheries productivity.    
3. Implementation of syncronized closures can outperform other intervnetions, although there are several factors to consider that are not included in this short analysis.    

References









